---
title: "assignment-b1-mbeletsky"
output: github_document
date: "2024-10-30"
---

# Setup

First, we'll load all the packages that we'll be using. 

```{r}
library(datateachr)
library(tidyverse)
library(testthat)
```

I've included "datateachr" so that we have access to the "vancouver_trees" dataset I'll be using as an example. We'll be using "tidyverse" to take advantage of the "dplyr" package. I've loaded "testthat" for the tests we're going to run later on the function created in this assignment. 

# Exercises 1 and 2: Make and Document a Function

I will make a function that bundles a specific **group_by()** %>% **summarize()** workflow.

I want to turn the following workflow into a function: counting how many unique observations there are for a variable.

```{r}
exercise1_example <- vancouver_trees %>%
  group_by(neighbourhood_name) %>%
  summarize(n_species = n_distinct(species_name))
print(exercise1_example)
```
Here, we have grouped the "vancouver_trees" dataset by neighbourhood and then summarized how many different species of tree there are in each neighbourhood. This workflow could also be useful for: counting how many genera there are per neighbourhood, or counting how many species there are in each genus.

Now, I will make a function that can accomplish this workflow. 

```{r}
#' @title Group by, then summarize unique observations
#' @description Group your dataset by one variable, then summarize counts of unique observations for another variable. Rows will be sorted from highest counts to lowest. 
#' @param {Object} data - A dataframe you are working with. I named this argument "data" to keep things simple and to the point, and for existing functions, it's common for the argument "data" to be a dataframe. 
#' @param {Object} group_var - The variable (column) you wish to group your data by. I named this argument "group_var" because it's the variable (var) you are wanting to group by (group) and to distinguish it easily from the other variable we are using as an argument. 
#' @param {Object} count_var - The variable (column) you wish to count unique observations for. I named this argument "count_var" because it's the variable (var) you are wanting to count unique observations in (count) and to distinguish it easily from the other variable we are using as an argument. 
#' @return Tibble that shows counts of unique observations within one variable, for each group you specify. 
group_then_sumz <- function(data, group_var, count_var) {
  table <- data %>%
    drop_na({{ group_var }}) %>%
    group_by({{ group_var }}) %>%
    drop_na({{ count_var }}) %>%
    summarize(n_distinct = n_distinct({{ count_var }})) %>%
    arrange(desc(n_distinct)) 
as_tibble(table)
print(table)
}
```

Here I've made a function called **group_then_sumz()** that takes in the arguments "data" (a dataframe), "group_var" (name of the column you want to group by), and "count_var" (name of the column you want to count unique observations for). It outputs a tibble that shows you unique counts for your variable of interest grouped by your other variable of interest, and sorts the counts from high to low.

I've dealt with potential NA values by using **drop_na()** on both the variable we want to group by and the one we want to summarize unique observations for. I've also included **arrange()** and used **desc()** to sort the values of the counted unique observations from high to low, so that the table is organized in a useful way. 

I've put the workflow itself into a new object created by the function, "table" which I coerce into a tibble with **as_tibble** at the end and then display the table with our results on the screen using **print()**.

# Exercise 3: Examples

## Group by neighbourhood, count unique species

It's time to show this function in action. Let's begin by grouping by neighbourhood and counting unique species. 

```{r}
group_then_sumz(vancouver_trees, neighbourhood_name, species_name)
```
Here we have a nice table showing how many distinct tree species there are in each Vancouver neighbourhood. We find out that Hastings-Sunrise, Kitsilano, and Renfrew-Collingwood have the most unique tree species. 

## Group by genus, count unique species

For the next example, let's group by genus and count how many distinct tree species there are in Vancouver that belong to each genus. 

```{r}
group_then_sumz(vancouver_trees, genus_name, species_name)
```
Now we have a nice table showing how many distinct tree species there are in Vancouver in each genus. The genera with the most unique species found in Vancouver are _Acer_, _Prunus_, and _Magnolia_.

## Group by genera, count unique cultivars

For this example, we will group by genus again and count how many cultivars there are in Vancouver in each genus.

```{r}
group_then_sumz(vancouver_trees, genus_name, cultivar_name)
```
Now that we've grouped by genus and summarized unique cultivars, we see that _Acer_, _Prunus_, and _Magnolia_ have the most unique cultivars within them. 

# Exercise 4: Test the Function

Now, we will make sure the function is working as it's supposed to by running tests. 

## Test 1: expect_is()

Within our function, we print out a tibble with the results which should also be a function because the output depends on the input data and variables. We can test if the function **group_then_sumz()** and the output object "table" are indeed functions by using **expect_is()** and using the arguments object and "function". 

```{r}
test_that("group_then_sumz and table are functions", {
expect_is(group_then_sumz, "function")
expect_is(table, "function")
})
```

The test passes, so we have confirmed that **group_then_sumz()** and "table" are both functions. 

## Test 2: expect_no_error()

We can use the test function **expect_no_error()** to check that with several non-redundant vector types, we don't run into any error messages.  

```{r}

test_that("different input variables produce no errors", {

# Vectors with no NA's
expect_no_error(group_then_sumz(vancouver_trees, neighbourhood_name, species_name), message = NULL, class = NULL)

# Vector with NA's
expect_no_error(group_then_sumz(vancouver_trees, genus_name, cultivar_name), message = NULL, class = NULL)

# Vector of different type (date)
expect_no_error(group_then_sumz(vancouver_trees, neighbourhood_name, date_planted), message = NULL, class = NULL)

# Vector of different type (numeric)
expect_no_error(group_then_sumz(vancouver_trees, neighbourhood_name, on_street_block), message = NULL, class = NULL)

})
```

I've tested the function with four different inputs: vectors with no NA's (neighbourhood_name and species_name), vector with NA's (cultivar_name), a date vector (date_planted), and a numeric vector (on_street_block). The test passed, meaning none of these input variables that we might be interested in generate an error when the function is run. 

## Test 3: expect_equal()

We can use the test function **expect_equal()** to test that our new function **group_then_sumz()** creates an output that is equivalent to running the **group_by()** %>% **summarize()** workflow manually. 

```{r}
test3_manual <- vancouver_trees %>%
  drop_na(neighbourhood_name) %>%
  group_by(neighbourhood_name) %>%
  drop_na(species_name) %>%
  summarize(n_distinct = n_distinct(species_name)) %>%
  arrange(desc(n_distinct))

test3_function <- group_then_sumz(vancouver_trees, neighbourhood_name, species_name)
```

To test this, I will first store the manual workflow in a new object called "test3_manual". I'll continue using the "vancouver_trees" dataset as an example and I'll group by neighbourhood and summarize number of distinct species. Then, I'll store the function that is intended to do the same thing in the new object "test3_function". I'll use vancouver_trees, neighbourhood_name, and species_name as the arguments as they correspond to the dataset we're using and the variables we're trying to group by and then summarize. 

Now, we can put "test3_function" (the new object) and "test3_manual" (the expected object) as arguments into the **expect_equal()** function within **test_that()**. **Expect_equal()** compares a computation to a reference value. We are expecting "test3_function" to match the reference, "test3_manual".

```{r}
test_that("running the function is equivalent to the manual workflow", {
expect_equal(test3_function, test3_manual)
})
```

The test is passed, indicating that using the function produces the same result as the manual workflow.
